### 2020-04-29 Tables/Table 1.Rmd
Created new figure 1 and table 1
shows smoothed and not smoothed ROC plots and AUC-ROC scores for the base specification

---
title: "Replication of Figure 1 and Table 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

library("kableExtra")
library("gridExtra")
library("tidyverse")
library("yardstick")
library("here")
```

### Introduction^[RM: As a way to maintain transparency and as a poor-mans version of preregistration, I'm pushing each step to GitHub, where it will have a time-stamp and track all changes.]

One issue that stood out as we read through Blair and Sambanis' (2020) article was how smooth their Receiver Operating Characteristic curves where in Figure 1 on page 14. A Receiver Operating Characteristic curve displays the balance between the true positive rate and the false positive rate, across a range of discrimination thresholds -- the value in which a zero to one probability is assumed to be an "onset" (here: civil conflict). Given that the DV is civil conflict onset (1/0) and that this is a rare event, these curves should resemble a step-like function. Looking through the Blair/Sambanis replication code, we noticed that they we using a rarely used in practice smoothing function when producing their Receiver Operating Characteristic curves.

This is important because Blair and Sambanis use a common performance metric -- Area Under the Receiver Operating Characteristic Curves (AUC-ROC) -- as evidence that their escalation model out preforms all others. However, by smoothing the Receiver Operating Characteristic Curves, they are misrepresenting the true AUC-ROC scores of their various models. Not only does this make model-to-model comparison difficult, it also makes it impossible to assess whether any of their models are performing better than random chance (a AUC-ROC score greater than 0.5) or whether their AUC-ROC scores surpass the (soft) industry standard that suggests that a model is performing well (an AUC-ROC score greater than 0.8).

To assess whether their decision to implement a smoothing function on the Receiver Operating Characteristic Curves for their various models impacts their results (namely, that the escalation model is superior to all others), we replicate Figure 1. However, rather than a two panel figure (one ROC plot their one-month-ahead forecasts and one for their six-months-ahead forecasts), we now include for panels (one panel for the two forecast windows with the smoothed curve (as they produce) and one panel for the two forecast windows that are not smoothed -- the industry standard).

Using these smoothed and non-smoothed curves, we then replicate Table 1 (page 13), which reports the AUC-ROC scores for these four different Receiver Operating Characteristic Curves. If the non-smoothed AUC-ROC scores are drastically different from those of the smoothed AUC-ROC score (a difference of &ge; 0.05) or if the non-smoothed AUC-ROC scores are < 0.8, would suggest that their decision to smooth-out the curves was improper. Further, if the the non-smoothed AUC-ROC score of the Escalation model is smaller or the same as of the non-smoothed AUC-ROC scores other models (Quad, Goldstein, and CAMEO) would suggests that the evidence presented in the paper is not robust to standard AUC-ROC calculation procedures.

Note: we are only reproducing the "Base specification" scores found in Table 1, as these are the what Blair and Sambanis use to support their claims. In order to ensure that this replication is a close as possible to the authors' original work, I will be using the replication code they provided at https://journals.sagepub.com/doi/suppl/10.1177/0022002720918923. There is a lot going on in their code, so I will be commenting out all elements that are not necessary for this replication process. Further, I will note any changes that I make this parsed code (i.e., removing the smoothing function). The raw replication files can be found at `GitHub/Blair-Sambanis-replication/Replication files/replication-3`, while all my edited code can be found at `GitHub/Blair-Sambanis-replication/Replication files/replication-3_BM`