### 2020-05-05 Tables/replicated-table-1.tex
Added stuff on table 1 replication

\begin{table}

\caption{\label{tab:tab-1}Replication of Table 1 in Blair and Sambanis (2020) with and without ROC smoothers}
\centering
\begin{tabular}[t]{ll>{\bfseries}lccccc}
\toprule
 & Horizon & ROC Smoothed & Escalation & Quad & Goldstein & CAMEO & Avg\\
\midrule
\addlinespace[0.3em]
\multicolumn{8}{l}{\textbf{Base Specification}}\\
\addlinespace[0.2em]
\hspace{1em} & 1 Month & Yes & 0.85 & 0.80 & 0.79 & 0.84 & 0.82\\
\cmidrule{3-8}
\hspace{1em}\hspace{1em} &  & No & 0.79 & 0.79 & 0.79 & 0.81 & 0.82\\
\cmidrule{2-8}
\hspace{1em} & 6 Months & Yes & 0.85 & 0.80 & 0.79 & 0.84 & 0.82\\
\cmidrule{3-8}
 &  & No & 0.79 & 0.79 & 0.79 & 0.81 & 0.82\\
\bottomrule
\end{tabular}
\end{table}