### 2020-06-08 table4/README.Rmd
fix up table 4 note

---
title: "Table 4 investigation"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readstata13)
library(ggplot2)
library(tidyr)
library(readr)
```

Is B&S Table 4 genered by scoring the civil war onset forecasts using civil war incidence?

The data used to train the forecasting models is `6mo_data_OOS.dta`. This just appears to be an extension of the regular 6 month training data that includes and additional

```{r}
data_6month_oos <- read.dta13("data/6mo_data_OOS.dta")

data_6month_oos %>% filter(year > 2013) %>% count(year, period)
```

The number of positive events--civil war onsets--matches what's in the training data, namely 20 onsets. There are no onsets recorded for 2016.

```{r}
data_6month_oos %>% count(incidence_civil_ns_plus1)
data_6month_oos %>% count(incidence_civil_ns ) 
data_6month_oos %>% filter(year==2016) %>% count(incidence_civil_ns)
```

The predictions seem to come from `6mo_predictions_escalation_OOS.dta`.

```{r}
preds <- read.dta13("data/6mo_predictions_escalation_OOS.dta") %>%
  as_tibble() %>%
  # for some reason the data vectors as character, not numeric. 
  mutate(across(-country, as.numeric))

preds
```

This actually has OOS predictions from 2008 to 2016 (this also means the latest forecasts for 2016 use stale 2007 data...):

```{r}
range(preds$year)
```

There are no onsets recorded in this either. But there are some missing values in the non-"plus1" version of the outcome.

```{r}
fcast2016 <- preds %>%
  filter(year==2016)

fcast2016 %>% count(incidence_civil_ns_plus1)
fcast2016 %>% count(incidence_civil_ns )

fcast2016 %>%
  dplyr::filter(is.na(incidence_civil_ns)) %>%
  dplyr::select(country, year)
```

If you start out with an incidence civil war variable, like the name here implies, and you convert it to onset, it might be that this was done by keeping the first year as a "1" and setting the rest to missing. Does it look like this was done here?

```{r}
data_6month_oos %>%
  select(year, period, country_iso3, incidence_civil_ns, incidence_civil_ns_plus1) %>%
  group_by(country_iso3) %>%
  mutate(any_non0 = any(is.na(incidence_civil_ns)) | any(incidence_civil_ns==1)) %>%
  ungroup() %>%
  filter(any_non0==TRUE) %>%
  mutate(country_iso3 = factor(country_iso3, levels = rev(unique(country_iso3))),
         date = sprintf("%s-%02s-01", year, ifelse(period %% 2 == 1, 1, 7)),
         date = as.Date(date)) %>%
  ggplot(aes(x = date, y = factor(country_iso3), 
             fill = factor(incidence_civil_ns))) +
  geom_tile() +
  scale_x_date(expand = c(0, 0)) +
  theme_bw()
```

This plot shows country-years--with countries on the *y*-axis and years on the *x*-axis. The fill color corresponds to whether the non-lagged outcome variable was coded as 0, 1, or missing.

It seems that all sequences of NA's that start after 2001 are preceded by a "1" coding. There are some isolated single half-year 1's, and Yemen has 2 "1"'s not separated by a non-civil war year.

Looking through the appendix:

- Cote d'Ivoire (CIV) indeed has a single-year civil war (p. 30)
- Libya also looks like it has a short civil war folled by longer one (p. 29)
- Yemen, ok, there are two long periods of civil war separated by a small gap; that probably explains the onset immediately after NA half year (p. 28)

Ok, so this indeed seems to be onset coded from an original incidence variable, as the name suggests anyways.

How do we get from that to B&S Table 4?

Aha, the culprit is in `6mo_make_confusion_matrix.do`. The script loads `predictions/6mo_predictions_escalation_OOS.dta`, drops everything prior to 2016. Then, on lines 50-52:

```stata
* Recode predictand

replace incidence_civil_ns=1 if incidence_civil_ns==.
```

So it seems that `incidence_civil_ns` went from some original incidence version to the onset + NA for ongoing coding in the data that is used for the R portion of the work; then prior to creating the confusion matrices for the forecast scoring, the onset + NA for ongoing coding is reversed, giving us back incidence coding. So indeed, the forecasts are scored using incidence.

Here's the alternative coding for the bottom portion, btw:

```stata
		  * Generate alternate coding of predictand

			gen incidence_civil_ns_alt=incidence_civil_ns
				replace incidence_civil_ns_alt=0 if country_name=="Colombia"
				replace incidence_civil_ns_alt=1 if country_name=="Turkey"
				replace incidence_civil_ns_alt=1 if country_name=="Burundi"
```

Redo a correct scoring against civil war onset.

```{r}
# Hand-code the Table 3 predictions;
# The predictions don't exactly match what is in Table 3, possibly because I
# only ran the 6 month OOS portion of +master.R after setting the seed at the
# top.
pred1 <- c("Nigeria", "India", "Iraq", "Somalia", "Syria", "Pakistan",
           "Philippines", "Turkey", "Afghanistan", "Russia", "Burundi", "Egypt",
           "Yemen", "Colombia", "Mali", "China", "Indonesia", "Ukraine",
           "Sudan", "Lebanon", "Thailand", "Iran", "Myanmar", "Montenegro",
           "Bangladesh", "Niger", "El Salvador", "France", "Ghana", "Tajikistan")
codes <- countrycode::countrycode(pred1, "country.name", "iso3c")
fcast2016$label <- as.integer(fcast2016$country %in% codes)
# should be 30
sum(fcast2016$label)

# What is the truth data in preds?
fcast2016 %>% count(incidence_civil_ns)
fcast2016 %>% count(incidence_civil_ns_plus1)

# Does this match the OOS data?
data_6month_oos %>% filter(period==31) %>% count(incidence_civil_ns)
data_6month_oos %>% filter(period==31) %>% count(incidence_civil_ns_alt1_plus1)

tab4_top <- fcast2016 %>%
  rename(Observed = incidence_civil_ns, Predicted = label) %>%
  replace_na(list(Observed = 0L)) %>%
  mutate(Observed = factor(Observed, levels = c("0", "1")),
         Predicted = factor(Predicted, levels = c("0", "1"))) %>%
  group_by(Observed, Predicted) %>%
  dplyr::summarize(n = n()) %>%
  ungroup() %>%
  tidyr::complete(Observed, Predicted, fill = list(n = 0)) %>%
  mutate(header = "Assuming Persistence")

# From 6mo_make_confusion_matrix.do, for the bottom of Table 3:
# replace incidence_civil_ns_alt=0 if country_name=="Colombia"
# replace incidence_civil_ns_alt=1 if country_name=="Turkey"
# replace incidence_civil_ns_alt=1 if country_name=="Burundi"
fcast2016 %>%
  filter(country %in% c("COL", "TUR", "BDI"))

data_6month_oos %>%
  filter(country_iso3 %in% c("COL", "TUR", "BDI")) %>%
  dplyr::select(country_iso3, year, incidence_civil_ns, incidence_civil_ns_plus1)

# Burundi has no ongoing conflict, so this is new onset
# Colombia has ongong conflict, so no difference
# Turkey has no ongoing conflict, so this is new onset
fcast2016_v2 <- fcast2016 %>%
  mutate(incidence_civil_ns = case_when(
    country=="TUR" ~ 1,
    country=="BDI" ~ 1,
    TRUE ~ incidence_civil_ns
  ))

tab4_bottom <- fcast2016_v2 %>%
  rename(Observed = incidence_civil_ns, Predicted = label) %>%
  replace_na(list(Observed = 0L)) %>%
  mutate(Observed = factor(Observed, levels = c("0", "1")),
         Predicted = factor(Predicted, levels = c("0", "1"))) %>%
  group_by(Observed, Predicted) %>%
  dplyr::summarize(n = n()) %>%
  ungroup() %>%
  tidyr::complete(Observed, Predicted, fill = list(n = 0)) %>%
  mutate(header = "Assuming Change")

tab4 <- bind_rows(tab4_top, tab4_bottom) %>%
  dplyr::select(header, everything()) %>%
  pivot_wider(names_from = "Predicted", values_from = n)

write_csv(tab4, "output/table4-fixed.csv")

tab4
```