### 2020-09-17 journal-survey/README.Rmd

---
title: "Journal survey"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Survey journals to check use of ROC/AUC and specifically how common it is to use smoothed ROC curves or AUCs derived from them.

"template.csv" is a template for adding a new journal issue (or other group of articles).

## Results

_(This README page is generated from a Rmarkdown file; re-knit that to update the results.)_

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
})

setwd(here::here("journal-survey"))

articles <- tibble(file = dir("data-raw", full.names = TRUE)) %>%
  mutate(data = suppressMessages(purrr::map(file, read_csv))) %>%
  unnest(data)

# mark articles that were referenced in B&S
articles$bsref <- as.integer(str_detect(articles$file, "bs\\-references"))
articles$file <- NULL

# take out items that are not coded yet (missing binary_outcome column)
articles <- articles[!is.na(articles$binary_outcome), ]

# if one of the payload columns is NA, set it to 0
articles$uses_auc[is.na(articles$uses_auc)] <- 0L
articles$uses_roc[is.na(articles$uses_roc)] <- 0L
articles$uses_smoothed_auc[is.na(articles$uses_smoothed_auc)] <- 0L
articles$uses_smoothed_roc[is.na(articles$uses_smoothed_roc)] <- 0L

cat("How many samples articles had binary outcome?\n")
table(articles$binary_outcome)

# The rest below is only for binary outcome articles
binary <- articles %>% filter(binary_outcome==1L)

smry <- binary %>%
  mutate(any_smoothed = uses_smoothed_auc | uses_smoothed_roc,
         any_empirical = uses_auc | uses_roc,
         uses_any = any_smoothed | any_empirical,
         only_smoothed = any_smoothed & !any_empirical,
         only_empirical = any_empirical & !any_smoothed,
         both = any_smoothed & any_empirical) %>%
  summarize(
    articles_coded = nrow(articles),
    articles_with_binary_outcome = n(),
    use_any_roc_measure = sum(uses_any),
    only_smoothed = sum(only_smoothed),
    only_empirical = sum(only_empirical),
    both = sum(both)
  ) %>%
  pivot_longer(everything())

knitr::kable(smry)

```