### 2020-09-18 journal-survey/README.Rmd
finish data entry for B&S references survey

---
title: "Journal survey"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Survey journals to check use of ROC/AUC and specifically how common it is to use smoothed ROC curves or AUCs derived from them.

"template.csv" is a template for adding a new journal issue (or other group of articles).

The sample consists of:

1. All references in the Blair and Sambanis paper, excluding books and non-academic articles. 
2. Two issues of JCR; 2020 volumes 1 and 9.

## Results

_(This README page is generated from a Rmarkdown file; re-knit that to update the results.)_

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
})

setwd(here::here("journal-survey"))

articles <- tibble(file = dir("data-raw", full.names = TRUE)) %>%
  mutate(data = suppressMessages(purrr::map(file, read_csv))) %>%
  unnest(data)

# mark articles that were referenced in B&S
articles$bsref <- as.integer(str_detect(articles$file, "bs\\-references"))
articles$file <- NULL

# take out items that are not coded yet (missing "journal" column)
articles <- articles[!is.na(articles$journal), ]

# if one of the payload columns is NA, set it to 0
articles$uses_auc[is.na(articles$uses_auc)] <- 0L
articles$uses_roc[is.na(articles$uses_roc)] <- 0L
articles$uses_smoothed_auc[is.na(articles$uses_smoothed_auc)] <- 0L
articles$uses_smoothed_roc[is.na(articles$uses_smoothed_roc)] <- 0L
articles$unclear_auc[is.na(articles$unclear_auc)] <- 0L
articles$unclear_roc[is.na(articles$unclear_roc)] <- 0L

cat("How many samples articles had binary outcome?\n")
table(articles$binary_outcome)

# The rest below is only for binary outcome articles
binary <- articles %>% filter(binary_outcome==1L)

binary <- binary %>%
  mutate(any_smoothed = uses_smoothed_auc | uses_smoothed_roc,
         any_empirical = uses_auc | uses_roc,
         any_unclear = unclear_roc | unclear_auc,
         uses_any = any_smoothed | any_empirical | any_unclear,
         only_smoothed = any_smoothed & !any_empirical & !any_unclear,
         only_empirical = any_empirical & !any_smoothed & !any_unclear,
         only_unclear = any_unclear & !any_empirical & !any_smoothed)

smry <- binary %>%
  summarize(
    items_coded = nrow(articles),
    articles_with_binary_outcome = n(),
    use_any_roc_measure = sum(uses_any),
    only_smoothed = sum(only_smoothed),
    only_empirical = sum(only_empirical),
    only_unclear = sum(only_unclear)
  ) %>%
  pivot_longer(everything())

knitr::kable(smry)

```

What items had unclear or ambiguous codings?

```{r}
binary %>% 
  filter(any_unclear) %>%
  select(title, year, journal, notes) %>%
  knitr::kable()
```

Note that both articles include Blair as co-author.

## Data

- "uses_roc", "uses_smoothed_roc", "unclear_roc": Does the paper include figures showing (empirical / smoothed / unclear) ROC curves?
- "uses_auc", "uses_smoothed_auc", "unclear_auc": Does the paper include tables showing (empirical / smoothed / unclear) AUC-ROC values?

If the paper includes visually obvious empirical ROC curves, we assumed that any corresponding AUC-ROC values were also generated from empirical ROC curves.

Some papers had AUC values but no ROC curves. In these cases we consulted replication code. These are listed below.

```{r}
articles %>% 
  filter(!uses_roc & uses_auc) %>%
  select(title, year, journal, notes) %>%
  knitr::kable()
```

```{r}
write_csv(articles, "coded-items.csv")
```